\large UNet-CRF Integration for Sequence Labeling on EEG Data

### 2024-12-09 manuscript/slides_results.tex
Add files via upload

\usepackage[linesnumbered,algoruled,boxed,lined]{algorithm2e}
\usepackage{amsfonts,amsmath,amssymb}
\usepackage{bbm}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{color}
\usepackage{graphicx}
\graphicspath{{./}{../image/}} % graphic path
\usepackage{hyperref}
\hypersetup{colorlinks,citecolor=blue,linkcolor=[RGB]{50,50,172}}
\usepackage{multicol}
\usepackage{multirow}
\usepackage[authoryear]{natbib}
\usepackage{setspace}
\usepackage{soul}
\usepackage{textpos}
\usepackage{tikz}

%% notations; only define those you use frequently
\newcommand{\EE}{{\mathbb{E}}}
\newcommand{\PP}{{\mathbb{P}}}
\newcommand{\QQ}{{\mathbb{Q}}}
\newcommand{\Fb}{\mathbf{F}}
\newcommand{\Ib}{\mathbf{I}}
\newcommand{\hb}{\mathbf{h}}
\newcommand{\xb}{\mathbf{x}}
\newcommand{\one}{\mathbbm{1}}
\newcommand{\Ocal}{\mathcal{O}}

@@

\title[\textcolor{black}{UNet-HCRF\_EEG}]{\large
UNet-HCRF Integration for Sequence Labeling on EEG Data}

@@

\begin{frame}{Introduction}
\begin{itemize}
\setlength{\itemsep}{1.5em}
\item Electroencephalography (EEG) pattern recognition tasks are crucial for
healthcare, where accurate and timely data classification can significantly
impact patient outcomes.
\item Deep learning has recently emerged as a powerful tool for sequence
labeling.
\item Conditional random fields (CRFs) have been employed for modeling label
dependencies, demonstrating promising results.
\end{itemize}
\end{frame}

\begin{frame}{Framework}
\begin{figure}[tbp]
\centering
\includegraphics[width=\textwidth]{model}
\end{figure}
\end{frame}

@@

\begin{frame}{Variational Inference of HCRF}
\begin{small}
\begin{itemize}
\item The mean-field varepsilon family approximation of
$\log P(\hb \mid y,\xb,\theta)$ is expressed as
$\QQ(h) = \prod_{i=1}^N q_i(h_i)$.
\item The evidence lower bound (ELBO) under the mean-field family is:
\begin{equation*}
\begin{split}
& \mathrm{ELBO}(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \vert \xb; \theta)
- \EE_{\QQ(\hb)} \log q(\hb) \\
&\quad = \sum_{i=1}^N \EE_{q_i(h_i)}
\left[ \phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) \right] \\
&\quad + \sum_{i=1}^N \sum_{j=1}^N \sum_{l,l^\prime} \one_{\{i \neq j\}} 
\QQ(h_i = l) \QQ(h_j = l^\prime) \mu(l, l^\prime) \\
&\quad \left[ \omega_1 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
\right) + \omega_2 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\gamma^2}
\right) \right] \\
&\quad - \sum_{i=1}^N \EE_{q_i} \log q_i(h_i) + \vartheta(y, \xb; 
\varpi).
\end{split}
\end{equation*}
\end{itemize}
\end{small}
\end{frame}

\begin{frame}{Variational Inference of HCRF (cont.)}
\begin{itemize}
\item By the coordinate ascent variational inference (CAVI) algorithm, fixing
$q_j(h_j), \forall j \neq i$, the optimal $q_i(h_i)$ that maximizes
$\mathrm{ELBO}(q_i(h_i))$ is:
\begin{equation*}
\QQ(h_i) \propto \exp\{\EE_{\QQ(\hb_{-i})}
\log\PP(y, h_i, \hb_{-i} \mid \xb; \theta)\}.
\end{equation*}
\item The explicit form of $q_i(l)$ is derived as follows:
\begin{equation*}
\begin{split}
& q_i(l) = \frac{1}{Z_i} \exp \Bigg\{
\phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta)
+ \sum_{j \neq i} \sum_{l^\prime} q(l^\prime) \mu(l, l^\prime) \\
& \left[ \omega_1 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
\right)
+ \omega_2 \exp \left(
- \frac{\left\lvert p_i - p_j 
\right\rvert^2}{2\eta_\gamma^2}
\right)
\right]
\Bigg\},
\end{split}
\end{equation*}
where $Z_i$ is the normalization constant.
\end{itemize}
\end{frame}

@@

\begin{frame}{Harmful Brain Activity Classification Dataset}
\begin{itemize}
\item The dataset contains 50-second long EEG samples covering a 10-minute
window as well as spectrograms matching each sample.
\bigskip
\item The EEG samples were labeled by expert annotators to determine which of
the six features each spectrogram belonged to.
\bigskip
\item The longer 10-minute window allows for the model to pick up on further
insights regarding the nature of the brain activity.
\end{itemize}
\end{frame}

@@

\begin{frame}{Make Spectrogram from EEG}
\begin{itemize}
\item Bipolar Montage combines multiple EEG electrode signals by calculating
the difference between adjacent electrodes, yielding signals representing
specific brain regions.
\end{itemize}
\begin{figure}[tbp]
\centering
\includegraphics[width=.7\textwidth]{inion}
\end{figure}
\end{frame}

@@

\begin{frame}{Make Spectrogram from EEG (cont..)}
\begin{figure}[tbp]
\centering
\vspace{-.5em}
\includegraphics[width=.5\textwidth]{EEG_Signal}
\end{figure}
\vspace{-1em}
\begin{figure}[tbp]
\centering
\includegraphics[width=.55\textwidth]{EEG_Spectrogram}
\end{figure}
\end{frame}

>>>>

\usepackage[linesnumbered,algoruled,boxed,lined]{algorithm2e}
\usepackage{amsfonts,amsmath,amssymb}
\usepackage{bbm}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{color}
\usepackage{graphicx}
\graphicspath{{./}{../image/}} % graphic path
\usepackage{hyperref}
\hypersetup{colorlinks,citecolor=blue,linkcolor=[RGB]{50,50,172}}
\usepackage{multicol}
\usepackage{multirow}
\usepackage[authoryear]{natbib}
\usepackage{setspace}
\usepackage{soul}
\usepackage{textpos}
\usepackage{tikz}
% \usepackage{algorithm}
% \usepackage{algpseudocode}
%% notations; only define those you use frequently
\newcommand{\EE}{{\mathbb{E}}}
\newcommand{\PP}{{\mathbb{P}}}
\newcommand{\QQ}{{\mathbb{Q}}}
\newcommand{\Fb}{\mathbf{F}}
\newcommand{\Ib}{\mathbf{I}}
\newcommand{\hb}{\mathbf{h}}
\newcommand{\xb}{\mathbf{x}}
\newcommand{\one}{\mathbbm{1}}
\newcommand{\Ocal}{\mathcal{O}}

@@

\title[\textcolor{black}{UNet-CRF\_EEG}]{\large
UNet-CRF Integration for Sequence Labeling on EEG Data}

@@

\begin{frame}{Introduction}
\begin{itemize}
\setlength{\itemsep}{1.5em}
\item Sequence labeling tasks are crucial for healthcare, where accurate and
timely data classification can significantly impact patient outcomes.
\item Deep learning has recently emerged as a powerful tool for sequence
labeling.
\item Conditional random fields (CRFs) have been employed for modeling label
dependencies, demonstrating promising results.
\end{itemize}
\end{frame}

@@

\begin{frame}{Variational Inference of HCRF}
\begin{small}
\begin{itemize}
\item The mean-field varepsilon family approximation of
$\log P(\hb \mid y,\xb,\theta)$ is expressed as
$\QQ(h) = \prod_{i=1}^N q_i(h_i)$.
\item The evidence lower bound (ELBO) under the mean-field family is:
\begin{equation*}
\begin{split}
& \textrm{ELBO}(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \vert \xb; \theta)
- \EE_{\QQ(\hb)} \log q(\hb) \\
&\quad = \sum_{i=1}^N \EE_{q_i(h_i)}
\left[ \phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) \right] \\
&\quad + \sum_{i=1}^N \sum_{j=1}^N \sum_{l,l^\prime} \one_{\{i \neq j\}} 
\QQ(h_i = l) \QQ(h_j = l^\prime) \mu(l, l^\prime) \\
&\quad \left[ \omega_1 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
\right) + \omega_2 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\gamma^2}
\right) \right] \\
&\quad - \sum_{i=1}^N \EE_{q_i} \log q_i(h_i) + \vartheta(y, \xb; 
\varpi).
\end{split}
\end{equation*}
\end{itemize}
\end{small}
\end{frame}

\begin{frame}{Variational Inference of HCRF (cont.)}
\begin{itemize}
\item By the coordinate ascent variational inference(CAVI) algorithm, fixing
$q_j(h_j), \forall j \neq i$, the optimal $q_i(h_i)$ that maximizes
$\mathrm{ELBO}(q_i(h_i))$ is:
\begin{equation*}
\QQ(h_i) \propto \exp\{\EE_{\QQ(\hb_{-i})}
\log\PP(y, h_i, \hb_{-i} \mid \xb; \theta)\}.
\end{equation*}
\item The explicit form of $q_i(l)$ is derived as follows:
\begin{equation*}
\begin{split}
& q_i(l) = \frac{1}{Z_i} \exp \Bigg\{
\phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta)
+ \sum_{j \neq i} \sum_{l^\prime} q(l^\prime) \mu(l, l^\prime) \\
& \left[ \omega_1 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
\right)
+ \omega_2 \exp \left(
- \frac{\left\lvert p_i - p_j 
\right\rvert^2}{2\eta_\gamma^2}
\right)
\right]
\Bigg\},
\end{split}
\end{equation*}
where $Z_i$ is the normalization constant.
\end{itemize}
\end{frame}

@@

\begin{frame}{Harmful Brain Activity Classification Dataset}
\begin{itemize}
\item The dataset contains 50-second long electroencephalography (EEG) samples
covering a 10-minute window as well as spectrograms matching each sample.
\bigskip
\item The EEG samples were labeled by expert annotators to determine which of
the six features each spectrogram belonged to.
\bigskip
\item The longer 10-minute window allows for the model to pick up on further
insights regarding the nature of the brain activity.
\end{itemize}
\end{frame}

@@

\begin{frame}{Make Spectrogram from EEG}
\begin{itemize}
\item Bipolar Montage combines multiple EEG electrode signals by calculating
the difference between adjacent electrodes, yielding signals representing
specific brain regions.
\end{itemize}
\begin{figure}[tbp]
\centering
% \includegraphics[width=.7\textwidth]{inion}
\end{figure}
\end{frame}

@@

\begin{frame}{Make Spectrogram from EEG (cont..)}
\begin{figure}[tbp]
\centering
\vspace{-.5em}
% \includegraphics[width=.5\textwidth]{EEG_Signal}
\end{figure}
\vspace{-1em}
\begin{figure}[tbp]
\centering
% \includegraphics[width=.55\textwidth]{EEG_Spectrogram}
\end{figure}
\end{frame}
\begin{frame}
	\frametitle{Mean Field Inference In Hidden CRF }
	\begin{algorithm}[H]
		\caption{Mean-field inference in fully-connected CRF}
		\KwIn{$Q_i(h_i) \gets U_i(h_i), \; i = 1, 2, \dots, N$ \tcp*[f]{Initialize $Q(h)$}}
		\While{not reach max iteration number}{
		    $\hat{Q}_i^{(m)}(h_i) \gets \sum_{j \neq i} k^{(m)}(f_i, f_j) Q_j(h_i)$ for all $m$ \tcp*[f]{Message Passing}\;
		    $\tilde{Q}_i^{(m)}(h_i) \gets \sum_{l \in \mathcal{L}} \mu^{(m)}(h_i, l) \sum_m \omega^{(m)} \hat{Q}_i^{(m)}(l)$ \tcp*[f]{Compatibility Transform}\;
		    $Q_i(h_i) \gets \exp \{\phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) - \tilde{Q}_i(x_i)\}$ \tcp*[f]{Local Update}\;
		    Normalize $Q_i(h_i)$\;
		}
		\end{algorithm}
\end{frame}

\begin{frame}
	\frametitle{Training Schedule}
	The loss function was designed as Kullback-leibler (kl) divergence.	
	Pre-train Unet learning rate schedule
	% \vspace{-1em}
	\begin{figure}[tbp]
	\centering
	\includegraphics[width=.55\textwidth]{pretrain}
	\end{figure}
	Fine tuning HCRF  learning rate schedule
	% \vspace{-1em}
	\begin{figure}[tbp]
	\centering
	\includegraphics[width=.55\textwidth]{finetune}
	\end{figure}
\end{frame}

\section{Results and Conclusion}

\begin{frame}
	\frametitle{Results}
	We evaluate our model by 5 fold cross-validation.
	\begin{itemize}
		\item Our model achieves a \textbf{0.4992} KL-divergence.
		\item The baseline model from Kaggle using EfficientNet obtains a 0.59 cross-validation KL-divergence.
	\end{itemize}
	However, there still are many teams on Kaggle achieve even lower KL-divergence(0.2-0.3) by using large model like Vision Transformers(ViT).
\end{frame}
\begin{frame}
	\frametitle{Conclusion}
	\begin{itemize}
		\item We build a Hidden Conditional Random Fields for EEG based Spectrograms image classification.
		\item The latent CRF layer is desgined to capture the latent segmentation information of the output of Neural Networks.
		\item A variational mean-field inference scheme is used to address the posterior distribution.
		\item We integrate the Unet and HiddenCRF into an end-to-end training structure. The Neural Networks feature learning and 
		variational inference can be accelerated jointly by GPUs.
	\end{itemize}
\end{frame}
